---
title: "Interpreting RTLbase outputs and visualizations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interpreting RTLbase outputs and visualizations}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

Each RTL algorithm returns a structured list with components that can be
inspected individually. This vignette highlights the most important fields and
how to translate them into diagnostics or figures.

## Algorithm 1: training diagnostics

```{r}
alg1$resamples <- lapply(alg1$results.all, function(res) res$train$overall)
alg1$resamples
```

`alg1$baselineSVM` is a matrix of hyperplane parameters (weights and bias) for
each source model. It is the foundation for Algorithms 2–6.

## Algorithm 2: robust moments

```{r}
alg2$U_robust       # robust mean of weights/bias
alg2$C_robust       # robust covariance matrix
alg2$U_robust_norm  # normalized hyperplane parameters
alg2$v_t            # orthonormal direction of change
```

High variance in `C_robust` indicates divergent source models that may benefit
from feature curation.

## Algorithm 3: shift compensation

```{r}
alg3  # vector of task-specific bias updates
```

Large magnitude shifts suggest strong batch effects between source and target
tasks. Consider narrowing `maximumLag` or focusing on a subset of features.

## Algorithm 4: bias update

```{r}
head(alg4$b_alg_norm)
```

This data frame stores the adjusted intercept for each task along with
intermediate density summaries (e.g., `c_j` counts and bandwidth estimates).

## Algorithm 6: normal vector rotation

```{r}
head(alg6$alg6_w_new)
```

Rows correspond to tasks, and columns correspond to updated weight vectors. Pair
these with `alg4$b_alg_norm` to build complete hyperplanes.

## Visualization helpers

`FinalViz()` produces ready-to-plot ggplot objects:

```{r}
viz$gg.figs$A     # combined accuracy/precision/recall plots across tasks
viz$comStats.sub  # tidy data frame of per-task summary metrics
```

- Use `ggsave()` to persist figures from `viz$gg.figs`.
- For per-task predictions, inspect `viz$ChangeAcross[[task_id]]$y_hat_alg6`
  alongside `viz$ChangeAcross[[task_id]]$y_test`.

## Exporting results

- Persist hyperplanes for reuse: `saveRDS(list(weights = alg6$alg6_w_new,
  bias = alg4$b_alg_norm), "hyperplanes.rds")`.
- Combine visualization summaries with sample metadata to track performance
  across experiments.
